---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# [`dv`](https://github.com/edwardlavender/dv): development tools for [`R`](https://www.r-project.org/)

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/dv)](https://CRAN.R-project.org/package=dv)

## Overview

[`dv`](https://github.com/edwardlavender/dv) is an [`R`](https://www.r-project.org/) package of development tools designed to support [RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html) and [`R`](https://www.r-project.org/) Package set up, development and maintenance. This package was inspired by the [`utils.add`](https://github.com/edwardlavender/utils.add) and [`gbclr`](https://github.com/cabi-org/gbclr) packages. Existing package functions are grouped into the following categories: 

* **Directory helpers**, including useful `here::here()` wrappers;
* **[RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html) tools**, including an [RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html) template generator;
* **[`R`](https://www.r-project.org/) package checks**, including  standard 'under-the-hood' checks for user inputs;

## Installation

1. **[`R`](https://www.r-project.org/) version.** [`dv`](https://github.com/edwardlavender/dv) requires [`R`](https://www.r-project.org/) version ≥ 4.1. You can check your current version with `R.version.string`. 

2. **Packages.** Installation requires the [`devtools`](https://github.com/r-lib/devtools) and [`pkgbuild`](https://github.com/r-lib/pkgbuild) packages, which can be installed with `install.packages(c("devtools", "pkgbuild"))`. 

3. **Rtools.** On Windows, package building requires [`Rtools`](https://cran.r-project.org/bin/windows/Rtools/). You can check whether [`Rtools`](https://cran.r-project.org/bin/windows/Rtools/) is installed with `pkgbuild::has_rtools()`. If `Rtools` is not installed, it is necessary to download and install the appropriate version of `Rtools` before proceeding by following the instructions [here](https://cran.r-project.org/bin/windows/Rtools/). 

4. **Authentication.** [`dv`](https://github.com/edwardlavender/dv) can be installed via [`devtools`](https://github.com/r-lib/devtools). Because [`dv`](https://github.com/edwardlavender/dv) is (at the time of writing) stored in a private repository, an authentication token is required for installation. Follow these steps to generate a token and save it in an environmental variable in [`R`](https://www.r-project.org/):

    * Go to https://github.com/settings/tokens and click 'Generate a new token'.
    * Select the 'Generate new token (classic)' option. If you are asked to confirm access at this stage, follow the on-screen instructions. You will arrive at the 'New personal access token (classic)' page. 
    * Fill in the 'Note' at the top of the page with a memorable comment that explains the purpose of the token (such as 'dv-installation'). 
    * In 'Expiration', the default settings are usually fine (but be aware that you may need to repeat this process to re-install an updated version of the package when the token expires). 
    * In 'Select scopes', select the 'repo' scope. 
    * Scroll to the bottom of the page and select 'Generate token'. 
    * You have now created a personal access token. 
    * Now go to [`R`](https://www.r-project.org/). If you are working on a [GitHub](https://github.com/) repository, make sure '.Rhistory' is registered in '.gitignore' so that the following steps are not tracked (run `usethis::git_vaccinate()` if necessary). Now save the token in an environmental variable called `GITHUB_PAT` via `Sys.setenv(GITHUB_PAT = "insert_copied_token_here")`. If you do this in an [`R`](https://www.r-project.org/) script (such as `01-secrets.R`) within a [GitHub](https://github.com/) repository, add the script to '.gitignore' so that you do not accidentally make your personal access token visible on [GitHub](https://github.com/).

5. **Installation.** Install [`dv`](https://github.com/edwardlavender/dv) along with its dependences and vignettes via [`devtools`](https://github.com/r-lib/devtools) as follows:

``` r
devtools::install_github("edwardlavender/dv", dependencies = TRUE, build_vignettes = TRUE)
```

## Functionality 

### Directory helpers 

**Directory helpers** facilitate specifying directories:

* `file_path()` and `here_path()` construct file paths;
* `here_*()` functions are `here::here()` wrappers for standard directories;
* `repair_path()` checks file/directory path validity;

### [RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html) tools

**[RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html) tools** support the organisation, maintenance and consistency of analytical pipelines. These tools include the following functions:

* `use_template_proj()` creates a template [RStudio](https://posit.co/) [Project](https://r4ds.had.co.nz/workflow-projects.html);
* `use_template_gitignore()` creates a template .gitignore file;
* `use_template_readme()` creates a template README file;
* `use_template_script()` creates a template script;
* `use_template_tree()` creates a directory tree;

### [`R`](https://www.r-project.org/) package checks (`check_*()`)

**[`R`](https://www.r-project.org/) package checks** include useful check functions for package functions:

* `check_class()` checks an object's class;
* `check_dir_exists()` checks a directory exists;
* `check_file_exists() checks a file exists;

## Usage 

```{r, eval = FALSE}
# Load and attach package
library(dv)
# Get help
help("dv")
# Get complete list of exported functions
help(package = "dv")
```

## Workflows 

[`dv`](https://github.com/edwardlavender/dv) encourages a standardised workflow for analytical work based on [RStudio](https://posit.co/) [Projects](https://r4ds.had.co.nz/workflow-projects.html). Package functions facilitate the creation of standardised projects, data acquisition, project management and replication. 

## Citation

Lavender, E. (2022). `dv`: Development tools for R. R package version 0.0.0.9000. https://github.com/edwardlavender/dv 

## Code of Conduct

Please note that the dv project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

---
